\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-27 metatheory-condensed.tex
Fixing typing rule in the main text
address a comment

\cref{fig:typing-selected} presents the syntax of \TT.
The new forms (relative to MLTT) are highlighted.

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
%
Formation rules for substitutions have\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) } the form 
$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
valid in the context~$\Dl$ yields terms valid in~$\Gm$. 
%
The two main forms of substitutions are weakening and extension:
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.

@@

\Rule[]
{}
{\goodWSig{\Gamma}{w\cdot}{0}}

@@

\Rule[name=WSig-Add]
  \quad \goodType{\Gamma}{A}{i}
  \quad \goodType{\Gamma, A}{B}{i}}
{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}

@@

\Rule[name=Ind-Univ]
{\goodWSig{\Gamma}{\tau}{n}}
{\goodType{\Gamma}{\bW \tau}{i}

@@

\Rule[name=Ind-Term]
{\goodTerm{\Gamma}{T}{\bW \tau}
  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)[p^1]}}
{\goodTerm{\Gamma}{"Wsup"~j~T~a~b}{El~("W"~T)} }

@@

\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}

@@

{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[p^2]))\; R[p^2]}{}}

@@

\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\tau}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
\quad j < N

@@

{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}

@@

\Rule[name=Rec-Constr]
{ 
  % \goodWSig{\Gamma}{\tau}{N}
\quad \goodTerm{\Gamma}{T}{\bW \tau}
% \quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}

@@

{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}

@@

\Rule[name=Rec-Beta]
{\goodTerm{\Gamma}{h}{\cL("RecSig"~\tau~R)} 
\\ \goodTerm{\Gamma}{T}{\bW\tau}}
{\goodTerm{\Gamma}{"Wrec"~h~("Wsup"~j~T~a~b) \equiv 
("app"~(("app"~("prjR"~j~h))[("id", a)]))[("id", "lam" ("Wrec"~h[p^1]~b))]

@@

% linkage, signature and compilation and seal
$$
\Rule[name=lsig/ept]
{""}
{\goodSig{\Gm}{\LSigEmp}{0}}
\quad
\Rule[name=ty/lkg]
{\goodSig{\Gm}{\lsig}{n}}
{\goodType{\Gm}{\TyLkg{\lsig}}{i}
\\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
\quad
\Rule[name=Lnkg Compile]
{ \goodTerm{\Gamma}{o}{\cL \sigma} 
}
{
  \goodTerm{\Gamma}{\Tkg {o}}{\cC \sigma}
}
$$
$$
\Rule[name=Sig Add]
{\goodSig{\Gamma}{\sigma}{n} 
 \quad \goodType{\Gamma}{A}{}
 \quad \goodSeal{\Gamma}{f}{\sigma}{A}
 \quad \goodType{\Gamma, A}{T}{i}}
{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
\quad
\Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[p^1]}} \Biggr)
$$
$$
\Rule[name=Ept Lnkg]
{}
{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
\quad
\Rule[name=Lnkg Add]
{ \goodTerm{\Gamma}{o}{\cL \sigma} 
\quad \goodType{\Gamma}{A}{}
\quad  {\goodSeal{\Gamma}{f}{\sigma}{A}}
 \quad \boxed{\goodTerm{\Gamma, A}{t}{T}}
}
{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
$$
$$
\Rule[name=lkg/proj]
{\goodTerm{\Gm}{o}{\TyLkg{\lsig}}}
{\goodTerm{\Gm}{\lkgproj{1} {o}}{\TyLkg {(\lsigproj{1} {\lsig})}}
\\ \boxed{\goodTerm{\Gm, \lsigprojT {\lsig}}{\lkgproj{2}{o}}{\lsigproj{2}{\lsig}}}
}
$$

>>>>

\cref{fig:typing-selected} presents selected syntax and typing rules of \TT.
The new forms (relative to MLTT) are highlighted.
We briefly review the base MLTT fragment first.

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
%
Typing judgments for substitutions have the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
valid in the context~$\Dl$ yields terms valid in~$\Gm$. 
%
The two main forms of substitutions are weakening and extension:
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.

@@

\Rule[name=wsig/ept]
{""}
{\goodWSig{\Gm}{\WSigEmp}{0}}

@@

\Rule[name=wsigeq/ept/eta]
{\goodWSig{\Gm}{\wsig}{0}}
{\goodWSig{\Gm}{\wsig \equiv \WSigEmp}{0}}

@@

\Rule[name=wsig/add]
{\goodWSig{\Gm}{\wsig}{n}
  \quad \goodType{\Gm}{A}{i}
  \quad \goodType{\Gm, A}{B}{i}}
{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}}
\Rule[name=wsig/universe]
{\goodWSig{\Gm}{\wsig}{n}}
{\goodType{\Gm}{\TyWSingle {\wsig}}{i}
\\ \goodTerm{\Gm}{\wsingle{\wsig}}{\TyWSingle{\wsig}}

@@

\Rule[name=ind/cstr]
{{\goodWSig{\Gm}{\wsig}{n}}
  \\ \goodTerm{\Gm}{a}{\wsigproj{j}{1}{\wsig}}
  \\ \goodTerm{\Gm, \sub{\wsigproj{j}{2}{\wsig}}{(id, a)}}{b}{\El{(\wcode{\wsig})}}}
{\goodTerm{\Gm}{\wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }

@@

\Rule[name=ty/casesig]
{\goodType{\Gm}{A}{}
\quad \goodType{\Gm, A}{B}{}
\quad \goodType{\Gm}{R}{}

@@

{\goodType{\Gm}{\CaseSig{A}{B}{R} \equiv (\TyPi {A} {(\TyPi {(\TyPi {B} {(\sub{R}{ \SubstComp {\pi_1}{\pi_1}})})} {\sub{R}{\SubstComp{\pi_1}{\pi_1}}})} )}{}}

@@

\Rule[name=reclkg/proj]
{ \goodWSig{\Gm}{\wsig}{N}
\\ \goodTerm{\Gm}{o}{\TyLkg{(\RecSig{\wsig}{R})}}
\\ j < N

@@

{\goodTerm{\Gm}{\Recproj{j}{o}}{\sub{(\CaseSig{(\pi_1^j \wsig)}{(\pi_2^j \wsig)}{R})}{\pi_1}}}

@@

\Rule[name=tm/wrec]
{ \goodWSig{\Gm}{\wsig}{N}
\\ \goodType{\Gm}{R}{}
\\ \goodTerm{\Gm}{r}{\TyLkg{(\RecSig{\wsig}{R})}}
\\ \goodTerm{\Gm}{w}{\El{(\wcode{\wsig})}}

@@

{\goodTerm{\Gm}{\Wrec{w}{r}}{R}}

@@

\Rule[name=tmeq/rec/beta]
{\goodTerm{\Gamma}{h}{\cL(\RecSig{\tau}{R})} 
\\ \goodTerm{\Gamma}{T}{\TyWSingle{\tau}}}
{\goodTerm{\Gamma}{\Wrec{h}{(\wsup{\wsig}{a}{b})} \equiv 
(\app{((\app{(\Recproj{j}{h})} )[("id", a)])} )[("id", \lam{(\Wrec{h[\pi_1]}{b})})]

@@

% linkage, signature and compilation and seal
$$
\Rule[name=lsig/ept]
{""}
{\goodSig{\Gm}{\LSigEmp}{0}}
\quad
\Rule[name=ty/lkg]
{\goodSig{\Gm}{\lsig}{n}}
{\goodType{\Gm}{\TyLkg{\lsig}}{i}
\\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
\quad
\Rule[name=lnkg/pkg]
{ \goodTerm{\Gm}{o}{\TyLkg {\lsig}} 
}
{
  \goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\lsig}}
}
$$
$$
\Rule[name=lsig/add]
{\goodSig{\Gm}{\lsig}{n} 
 \quad \goodType{\Gm}{A}{}
 \quad \goodSeal{\Gm}{\seal}{\lsig}{A}
 \quad \goodType{\Gm, A}{T}{i}}
{\goodSig{\Gm}{(\LSigAdd {\lsig}{\seal}{T})}{n+1}}
\quad
\Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[p^1]}} \Biggr)
$$
$$
\Rule[name=lnkg/ept]
{""}
{\goodTerm{\Gm}{\LkgEmp}{\TyLkg {\LSigEmp}}  
}
\quad
\Rule[name=lkg/add]
{ \goodTerm{\Gm}{o}{\TyLkg {\lsig}} 
\\ \goodType{\Gm}{A}{}
\\  \goodSeal{\Gm}{\seal}{\lsig}{A} 
\\  \boxed{\goodTerm{\Gm, A}{t}{T}}
}
{\goodTerm{\Gm}{(\LkgAdd {o} {t})}{\TyLkg{\LSigAdd {\lsig} {\seal} {T}}}
}
$$
$$
\Rule[name=lkg/proj]
{\goodTerm{\Gm}{o}{\TyLkg{\lsig}}}
{\goodTerm{\Gm}{\lkgproj{1} {o}}{\TyLkg {(\lsigproj{1} {\lsig})}}
\\ \boxed{\goodTerm{\Gm, \lsigprojT {\lsig}}{\lkgproj{2}{o}}{\lsigproj{2}{\lsig}}}
}
$$
